---
title: "Example Clustering of Google Location History"
author: "Dr. Gordon McDonald"
date: "24/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(stdbscanr)
library(lubridate)
library(leaflet)
```

Example to cluster a random day in Feb 2017 of my Google location history's trajectory, using stdbclust.

# Read location history

Read in my location history as a .csv, and set time column to be a datetime.
```{r}
location_data <- fread("location.csv")[
  ,time := ymd_hms(`time`)]
#order by time
setkey(location_data,time)
#calc time intervals between points
location_data[,time_inc := as.numeric(time - shift(time), 
                                      units = "mins")]
```

A small sample of the data looks like this:
```{r}
location_data[699:708,]
```

The data helpfully comes with labels attached by google but we won't look at these for clustering:
```{r}
unique(location_data[,.(name,category)])
```

# Get clusters

Get each visit to a place as a space-time cluster
```{r}
location_with_visits <- 
  get_clusters_from_data(df = location_data,
                         x = "longitude", 
                         y = "latitude", 
                         t = "time",
                         eps = 0.0005, #0.0005 latitude/longitude ~ 50m either way in sydney
                         eps_t = 10, #10 min
                         minpts = 10)

setnames(location_with_visits, old = "cluster", new = "visit") #rename cluster to visit
```

# Visit Summary

Make a summary table of visits
```{r}
#Define a mode function to get the most common label
mode <- function(x) { names(which.max(table(x))) }

#data.table summary table of visits
visits <-                                   # dplyr equivalents:
  location_with_visits[     
    !is.na(visit),                          # filter(!is.na(visit))
    .(n = .N,                               # summarise(n = n(),
      latitude = mean(latitude),            #          latitude = mean(latitude),
      longitude = mean(longitude),          #          longitude = mean(longitude),
      time_spent = sum(time_inc,na.rm = T), #          time_spent = sum(time_inc),
      google_label = mode(name)             #          google_label = mode(name))
    ),    
    by=visit]                               # group_by(visit)
```

Let's have a look at the clusters we've found:
```{r}
visits[,.(visit, google_label, latitude, longitude, time_spent)]
```
(and no, I don't live there any more!)

# Cluster visits together into places

To do this, we will abuse stdbclust to do dbclust on the visits, in order to group together visits that are in the same place. This requires creating a copy of longitude and stuffing it in instead of time.
```{r}
places <- 
  get_clusters_from_data(
    df = visits[,lon2:=longitude],
    x = "longitude", 
    y = "latitude", 
    t = "lon2",
    eps = 0.002, #0.002 latitude/longitude ~ 200m either way in sydney
    eps_t = 0.002, #0.002 latitude/longitude ~ 200m either way in sydney
    minpts = 1)[
      ,`:=`(place = cluster, # rename to place
            cluster = NULL)]

#make a dictionary to convert between visits and places
visits_to_places <- unique(na.omit(places[,.(visit, place)]))

#merge the places back on top of the location/trajectory data
location_with_visits_and_places <- 
  merge(location_with_visits,            # merge on the only column in common, `visit`
        visits_to_places, 
        all.x = T)[                      # left join
          ,`:=`(place = factor(place),   # set place and visit to be factors
                visit = factor(visit))]
```

# Make a summary of the places we've found

```{r}
#data.table summary table of places
places_summary <-
  location_with_visits_and_places[         # dplyr equivalents:
    !is.na(visit),                         #  filter(!is.na(visit))
    .(n = .N,                              #  summarise(n = n(),
      latitude = mean(latitude),           #            latitude = mean(latitude),
      longitude = mean(longitude),         #            longitude = mean(longitude),
      time_spent = sum(time_inc,na.rm = T),#            time_spent = sum(time_inc,na.rm = T),
      visits = length(unique(visit)),      #            visits = length(unique(visit)),
      google_label = mode(name)),          #            google_label = mode(name))
    by=place]                              # group_by(place)
```

So what does it look like?
```{r}
places_summary[,.(place, google_label, visits, latitude, longitude, time_spent)]
```
We can see that I made 4 visits home (or had GPS on on 4 separate occasions), and Coogee beach has been absorbed into Coogee Pavillion, becasue of the 200m radius in generating `places`.

# Plot trajectory and places on a leaflet map

```{r}
# order by time
setkey(location_with_visits_and_places, time)

# plot on leaflet map
leaflet(data = places_summary) %>% 
  addProviderTiles("OpenStreetMap.Mapnik") %>% 
  addMarkers(popup = ~paste0("Time spent: ",round(time_spent/60, 1), " hours.<br>",
                             "Location ",place,": ",google_label,"<br>",
                             "Visits: ", visits)) %>% 
  addPolylines(data = location_with_visits_and_places, 
               lat = ~latitude, 
               lng = ~longitude)
```

